---
title: "Comparative analysis of multiple datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=6, fig.height=6,fig.align = "center") 
```
This vignette will show how to use NeuronChat to perform comparative analysis on multiple datasets, by taking the comparison between VISp and ALM as an example. 

# Load NeuronChat library

```{r setup}
library(NeuronChat)
library(CellChat)
```

# Part I: Create NeuronChat object list

```{r}
data(list='cortex_data') # including `target_df` and `meta`
cortex_1K <- c('ALM','VISp')
subclass_CTX <- names(table(meta$subclass_label[meta$class_label=='Glutamatergic']))[grep('CTX',names(table(meta$subclass_label[meta$class_label=='Glutamatergic'])))] # subset the Glutamatergic subclasses
subclass_CTX
set.seed(1234)
cortex_list <- lapply(cortex_1K, function(y){
  region_name <- y; cell_class <- 'Glutamatergic';if(is.null(cell_class)){cell_class <- names(table(meta$class_label))}
  cell_idx <- which(meta$region_label %in% region_name & meta$class_label %in% cell_class & meta$subclass_label %in% subclass_CTX)
  target_df_single  <- target_df[cell_idx,]
  meta_tmp <- meta[cell_idx,];rownames(meta_tmp) <- meta_tmp$sample_name
  x <- createNeuronChat( t(as.matrix(target_df_single[,1:(dim(target_df_single)[2]-1)])),DB='mouse',group.by =   target_df_single$cell_subclass,meta=meta_tmp);
  x <- run_NeuronChat(x,M=100)
  return(x)
})
names(cortex_list) <- cortex_1K
```
```{r fig0, fig.height = 5, fig.width = 10}
# individually plotting the communication networks for VISp and ALM  
par(mfrow=c(1,2))
for(j in c(1,2)){
  netVisual_circle_neuron(cortex_list[[j]]@net$Glu_Grin3a, title.name = paste('Glu_Grin3a -- ',names(cortex_list)[j]),arrow.size = 0.5,margin=0.3,edge.width.max=8)
}
par(mfrow=c(1,2))
for(j in c(1,2)){
  net_aggregated_x <- net_aggregation(cortex_list[[j]]@net,method='weight')
  netVisual_circle_neuron(net_aggregated_x, title.name = names(cortex_list)[j],arrow.size = 0.5,margin=0.3,edge.width.max=8)
}
```

# Part II: Merge the NeuronChat object list 

```{r}
neuronchat_list <- mergeNeuronChat(cortex_list, add.names = names(cortex_list))
```

# Part III: Barplots to compare link count and weight of interaction pairs between ALM and VISp

```{r fig1, fig.height = 3, fig.width = 6}
# overall communication
p1 <- compareInteractions_Neuron(neuronchat_list,measure = c("weight"),comparison = c(1,2),group=c(1,2),show.legend = F)
p2 <- compareInteractions_Neuron(neuronchat_list,measure = c("count"),comparison = c(1,2),group=c(1,2),show.legend = F )
p1 + p2
```

```{r fig2, fig.height = 12, fig.width = 9}
# communication for individual interaction pairs
g1 <- rankNet_Neuron(neuronchat_list,mode='comparison',measure = c("count"),comparison = 1:2,do.stat = F,tol = 0.1,stacked = F,font.size = 11)
g2 <- rankNet_Neuron(neuronchat_list,mode='comparison',measure = c("weight"),comparison = 1:2,do.stat = F,tol = 0.1,stacked = F,font.size = 11) 
g1+g2
```

# Part IV: Shared and specific interaction patterns across ALM and VISp

```{r fig3, fig.height = 4, fig.width = 5}
# compute functional similarity 
neuronchat_list <- computeNetSimilarityPairwise_Neuron(neuronchat_list, slot.name = "net", type = "functional",comparison = c(1,2))
# manifold learning
neuronchat_list <- netEmbedding(neuronchat_list,slot.name = "net_analysis", type = "functional",comparison = c(1,2))
# clustering on interactions 
neuronchat_list <- netClustering(neuronchat_list, slot.name = "net_analysis", type = "functional",comparison = c(1,2),k = 5)
# visualization
netVisual_embeddingPairwise_Neuron(neuronchat_list, slot.name = "net_analysis", type = "functional", label.size = 3.5,comparison=c(1,2),pathway.remove.show = FALSE,pathway.labeled = F)
```

```{r fig31, fig.height = 7, fig.width = 12}
# visualization zoom in
netVisual_embeddingPairwiseZoomIn_Neuron(neuronchat_list, slot.name = "net_analysis", type = "functional", label.size = 5,comparison=c(1,2),nCol=3)
```

```{r fig4, fig.height = 3.5, fig.width = 20} 
# heatmap for each interaction pattern
net12 <- neuronchat_list@net[c(1,2)]
net1 <- net12[[1]];names(net1) <- paste(names(net1),'--ALM',sep='')
net2 <- net12[[2]];names(net2) <- paste(names(net2),'--VISp',sep='')
net12_list <- append(net1,net2)
interaction_group <- neuronchat_list@net_analysis$similarity$functional$group$`1-2`
hlist <- list();gb_heatmap <- list()
library(grid); library(ComplexHeatmap);grid.newpage(); x_seq <- c(0,0.2,0.4,0.6,0.8)
for(j in 1:length(sort(unique(interaction_group),decreasing = F))){
  net_aggregated_group2 <- net_aggregation(net12_list[names(interaction_group[interaction_group==j])],method = 'weight')
  library(RColorBrewer);col_map = brewer.pal(8,"YlOrBr");
  h <- Heatmap(net_aggregated_group2, name = "Weight",
                        col = col_map,
                        cluster_rows = FALSE,cluster_columns=FALSE,
                        row_names_side='left',column_names_side='bottom',
                        row_title='Sender',row_title_side='left',
                        row_title_gp = gpar(fontsize = 16),
                        column_title='Receiver',column_title_side = "bottom",column_title_gp = gpar(fontsize = 16),column_names_rot = 60)
  gb_heatmap[[j]] = grid.grabExpr(draw(h,column_title=paste('pattern cluster',j), padding = unit(c(2, 2, 2, 2), "mm")) )
  pushViewport(viewport(x = x_seq[j], y = 1, width = 0.19, height = 1, just = c("left", "top"),xscale = c(0, 1), yscale = c(0, 1)));grid.draw(gb_heatmap[[j]]);popViewport()
}
```

